\speculake: Hiding Malware in Speculative Execution

### 2018-08-07 paper/figures.tex
update captions and clarifications to implementation

\newcommand{\FigSpectreOne}{
\begin{figure}[t]
    \centering
    \includegraphics[width=0.9\linewidth]{figures/spectre_1_training.png}
    \caption{\textbf{Speculative buffer overflow warm-up}\,---\, %
    The direct branch predictor must
    be trained to expect that a branch will go a specific way before speculative
    buffer overflows can be used. We varied the number of times a branch was trained
    to be taken and observed the fraction of times we achieved a speculative buffer overflow
    execution immediately after (measured by observing if a speculatively-loaded value was present
    in cache avgeraged over 20,000 trials). We find that a branch must be trained in a
    direction hundreds of times before it can start to be used in a speculative buffer overflow.}
    \label{fig:spectre-one}
\end{figure}
}

\newcommand{\FigSpectreOneModel}{
\begin{figure}[ht]
    \centering
    \includegraphics[width=0.9\linewidth]{figures/spectre_1_model.png}
    \caption{\textbf{Speculative Buffer Overflow}\,---\, %
    Safe input values are stored into a local array (Step 1),
    training the branch predictor to bypass the bounds check when given
    unsafe input (Step 2). This results in a speculative array store which
    overflows the bounds and overwrite a return address (Step 3). The CPU's 
    store-to-load forwarding will forward the speculative overflow to the 
    return (Step 4), where it will steer speculative execution to the 
    attacker-specified speculative entry point (Step 5). Eventually, the 
    original bounds check is resolved and control flow is redirected to the 
    cache timing function (Step 6). }
    \label{fig:spectre-one-model}
\end{figure}
}

\newcommand{\FigSpecMeasureNP}{
\begin{figure*}[t]
     \centering
        \includegraphics[width=0.9\textwidth]{figures/Speculative_measurements_no-parity.png}
        \caption{\textbf{Speculative limits}\,---\, %
    To find the number of instructions we can execute speculatively, we placed a speculative load
    after an increasing number of instructions, and measured the fraction of times the load
    was subsequently in cache. However, more complex instructions
    have stricter constraints, generally corresponding to the number of $\mu$-ops that an instruction
    comprises. We find that we can execute up to 220 instructions 
    speculatively, which corresponds with the size of the CPU's reorder buffer (ROB). We measure 
    the ROB size by timing two cache accesses with a variable number of 
    instructions between them. The step in cycles around 220 instructions indicates that the 
    second cache access was pushed out of the ROB defining an upper limit on size.  }
    \label{fig:spec-capacity-np}
\end{figure*}
}

\newcommand{\FigSpecMeasureParity}{
\begin{figure}[b]
     \centering
        \includegraphics[width=1.05\linewidth]{figures/Speculative_measurements_parity.png}
        \caption{\textbf{Hyperthreading}\,---\, %
        The trigger program influences the payload program via the indirect branch predictor
        when running on the same logical CPU core. Hyperthreads also share an indirect branch
        predictor despite running on different logical cores. We measured the impact of
        same logical core (warm colors) vs paritiy hyperthreads (cool colors, denoted by (P))
        on the size of the speculative execution window. As hyperthreads run concurrently, they do not
        need to wait for an OS context switch between processes, but the two processes directly
        contend for ROB entries. As a result, we find that hyperthreads generally have a smaller
        speculative execution window.}
    \label{fig:spec-capacity-parity}
\end{figure}
}

@@

\newcommand{\FigGeneralModel}{
\begin{figure}[t]
    \centering
        \includegraphics[width=0.4\textwidth]{figures/general_model.pdf}
        \caption{\textbf{ExSpectre model}\,---\, %
        General model of speculative computation within the payload 
        process when triggered. The \textit{Speculative Gadget} has read-only
        access to all memory within the process, but can return updates/results via a
        cache side channel (via access to the \texttt{probe\_array}).
        The process can subsequently make
        \textit{Resultant Computations} based on the value returned from
        the cache side channel to update the state of the \textit{Real World} process.}
    \label{fig:general_model}
\end{figure}
}

\newcommand{\FigSpecBandwidth}{
\begin{figure}[t]
    \centering
        \includegraphics[width=0.5\textwidth]{figures/Speculative_bandwidth}
        \caption{\textbf{Speculative Bandwidth}\,---\, %
        Using the speculative primitive, 1KB of data
        can be decrypted and exfiltrated at a speed between 5 and 6 kbps from
        the speculative world with 20 iterations per round for guaranteed
        correctness.  Varying the width of the channel results in competing
        performance factors, with an optimal throughput achieved at 8~bits.}
    \label{fig:spec_bandwidth}
\end{figure}
}

\newcommand{\FigSpasmModel}{
\begin{figure}[b]
    \centering
        \includegraphics[width=0.46\textwidth]{figures/spasm_model.pdf}
        \caption{\textbf{SPASM model}\,---\, %
        Our SPASM emulator speculatively decrypts intructions, and emulates
        them in the real world. The \textit{Speculative Computation} decrypts
        the encrypted SPASM binary, returning the result through the side channel
        to allow the \textit{Resultant Computation} to update the emulated state
        (or make system calls) on behalf of the speculative world.}
    \label{fig:spasm_model}
\end{figure}
}

>>>>

\newcommand{\FigSpectreOne}{
\begin{figure}[t]
    \centering
    \includegraphics[width=0.9\linewidth]{figures/spectre_1_training.png}
    \caption{\textbf{Speculative buffer overflow warm-up}\,---\,%
    The direct branch predictor must
    be trained to expect that a branch will go a specific way before speculative
    buffer overflows can be used. We varied the number of times a branch was trained
    to be taken and observed the fraction of times we achieved a speculative buffer overflow
    execution immediately after (measured by observing if a speculatively-loaded value was present
    in cache avgeraged over 20,000 trials). We find that a branch must be trained in a
    direction hundreds of times before it can start to be used in a speculative buffer overflow.}
    \label{fig:spectre-one}
\end{figure}
}

\newcommand{\FigSpectreOneModel}{
\begin{figure}[ht]
    \centering
    \includegraphics[width=0.9\linewidth]{figures/spectre_1_model.png}
    \caption{\textbf{Speculative Buffer Overflow}\,---\,%
    Safe input values are stored into a local array (Step 1),
    training the branch predictor to bypass the bounds check when given
    unsafe input (Step 2). This results in a speculative array store which
    overflows the bounds and overwrite a return address (Step 3). The CPU's 
    store-to-load forwarding will forward the speculative overflow to the 
    return (Step 4), where it will steer speculative execution to the 
    attacker-specified speculative entry point (Step 5). Eventually, the 
    original bounds check is resolved and control flow is redirected to the 
    cache timing function (Step 6). }
    \label{fig:spectre-one-model}
\end{figure}
}

\newcommand{\FigSpecMeasureNP}{
\begin{figure*}[t]
     \centering
        \includegraphics[width=0.9\textwidth]{figures/Speculative_measurements_no-parity.png}
        \caption{\textbf{Speculative limits}\,---\,%
    To find the number of instructions we can execute speculatively, we placed a speculative load
    after an increasing number of instructions, and measured the fraction of times the load
    was subsequently in cache. However, more complex instructions
    have stricter constraints, generally corresponding to the number of $\mu$-ops that an instruction
    comprises. We find that we can execute up to 220 instructions 
    speculatively, which corresponds with the size of the CPU's reorder buffer (ROB). We measure 
    the ROB size by timing two cache accesses with a variable number of 
    instructions between them. The step in cycles around 220 instructions indicates that the 
    second cache access was pushed out of the ROB defining an upper limit on size.  }
    \label{fig:spec-capacity-np}
\end{figure*}
}

\newcommand{\FigSpecMeasureParity}{
\begin{figure}[b]
     \centering
        \includegraphics[width=1.05\linewidth]{figures/Speculative_measurements_parity.png}
        \caption{\textbf{Hyperthreading}\,---\,%
        The trigger program influences the payload program via the indirect branch predictor
        when running on the same logical CPU core. Hyperthreads also share an indirect branch
        predictor despite running on different logical cores. We measured the impact of
        same logical core (warm colors) vs paritiy hyperthreads (cool colors, denoted by (P))
        on the size of the speculative execution window. As hyperthreads run concurrently, they do not
        need to wait for an OS context switch between processes, but the two processes directly
        contend for ROB entries. As a result, we find that hyperthreads generally have a smaller
        speculative execution window.}
    \label{fig:spec-capacity-parity}
\end{figure}
}

@@

\newcommand{\FigGeneralModel}{
\begin{figure}[t]
    \centering
        \includegraphics[width=0.4\textwidth]{figures/general_model.pdf}
        \caption{\textbf{ExSpectre model}\,---\,%
        General model of speculative computation within the payload 
        process when triggered. The \textit{Speculative Gadget} has read-only
        access to all memory within the process, but can return updates/results via a
        cache side channel (via access to the \texttt{probe\_array}).
        The process can subsequently make
        \textit{Resultant Computations} based on the value returned from
        the cache side channel to update the state of the \textit{Real World} process.}
    \label{fig:general_model}
\end{figure}
}

\newcommand{\FigSpecBandwidth}{
\begin{figure}[t]
    \centering
        \includegraphics[width=0.5\textwidth]{figures/Speculative_bandwidth}
        \caption{\textbf{Speculative Bandwidth}\,---\,%
        Using the speculative primitive, 1KB of data
        can be decrypted and exfiltrated at a speed between 5 and 6 kbps from
        the speculative world with 20 iterations per round for guaranteed
        correctness.  Varying the width of the channel results in competing
        performance factors, with an optimal throughput achieved at 8~bits.}
    \label{fig:spec_bandwidth}
\end{figure}
}

\newcommand{\FigSpasmModel}{
\begin{figure}[b]
    \centering
        \includegraphics[width=0.46\textwidth]{figures/spasm_model.pdf}
        \caption{\textbf{SPASM model}\,---\,%
        Our SPASM emulator speculatively decrypts intructions, and emulates
        them in the real world. The \textit{Speculative Computation} decrypts
        the encrypted SPASM binary, returning the result through the side channel
        to allow the \textit{Resultant Computation} to update the emulated state
        (or make system calls) on behalf of the speculative world.}
    \label{fig:spasm_model}
\end{figure}
}

### 2018-08-07 paper/implementation.tex
comment out sbo impl for now
update captions and clarifications to implementation

We used 6 bits in the construction of this instruction set as our
goal is to limit the length of each opcode as much as possible. Note that this
is different from the goal in maximizing bandwidth, as our goal now is to
maximize instruction throughput. Given our short instructions, loading values into registers
requires shifting in 4-bits at a time. SPASM has two registers that act as a
pointer and working register, that can be used to perform jumps, arbitrary
memory reads and writes, and basic arithmatic. We also have a \texttt{syscall}
instruction that makes a real system call to the underlying operating system
with parameters loaded from the SPASM state, allowing us to interact with the
real world. Details of
the SPASM instruction set can be found in Appendix~\ref{appendix:spasm}.

In SPASM we have implemented multiple example programs that we encrypted and
loaded into a \speculake payload, which decrypts and emulates SPASM instructions
only when the corresponding trigger program is running.
We have implemented a \textit{HelloWorld}
program that prints to \texttt{stdout}, and a \texttt{FizzBuzz} program that
demonstrates control flow and arithmetic operations while printing to stdout.
Finally, we implemented a \textit{ReverseShell} program that opens and connects a TCP socket
to an attacker-chosen location
before executing a local shell and allowing the remote adversary
to issue shell commands on the victim machine.
Figure~\ref{fig:spasm_model} details the high-level flow of a SPASM payload.

@@

We discovered a candidate that corresponds to code in OpenSSL's
\texttt{nistp256.c} that contained 31 indirect jumps repeated 254 times each
handshake. This code is used during the TLS key exchange as the server computes
the ECDHE shared secret. We made a list of 31 source-destination address pairs
for these indirect jumps, and constructed a jump/ret chain to mimic the same
jump pattern in
our payload program. Our payload program mimicks the first 30 indirect jump
source/destination pairs, with a final jump going to a cache timing function
in our payload program. However, due to the prior pattern, this last jump is
frequently mis-speculated (about 3.5\% of the time), and instead goes to the
destination corresponding to the 31st jump in OpenSSL, which serves as our
speculative entry point.

@@

\subsection{Speculative Buffer Overflow Trigger}
\label{subsec:openssl-impl}

\FigSpectreOne

@@

We note significant unpredictability in the branch predictor, though we
show in Figure~\ref{fig:spectre-one} that a minimum around 200 safe input entries
to each dishonest entry are required before a distinguishable signal is present 
in the cache channel.

%%%%%%%%%%%%%%

>>>>

We used 6 bits in the construction of this instruction set as our goal is to
limit the length of each opcode as much as possible. Note that this is different
from the goal in maximizing bandwidth, as our goal now is to maximize
instruction throughput. Given our short instructions, loading values into
registers requires shifting in 4-bits at a time. SPASM has two registers that
act as a pointer and working register, that can be used to perform jumps,
arbitrary memory reads and writes, and basic arithmetic. We also have a
\texttt{syscall} instruction that makes a real system call to the underlying
operating system with parameters loaded from the SPASM state, allowing us to
interact with the real world. Details of the SPASM instruction set can be found
in Appendix~\ref{appendix:spasm}.

In SPASM we have implemented multiple example programs that we encrypted and
loaded into a \speculake payload, which decrypts and emulates SPASM instructions
only when the corresponding trigger program is running. We have implemented a
\textit{HelloWorld} program that prints to \texttt{stdout}, and a
\textit{FizzBuzz} program that demonstrates control flow and arithmetic
operations while printing to stdout. Finally, we implemented a
\textit{ReverseShell} program that opens and connects a TCP socket to an
attacker-chosen location before executing a local shell and allowing the remote
adversary to issue shell commands on the victim machine.
Figure~\ref{fig:spasm_model} details the high-level flow of a SPASM payload.

@@

We discovered a candidate that corresponds to code in OpenSSL's
\texttt{nistp256.c} that contained 31 indirect jumps repeated 254 times each
handshake. This code is used during the TLS key exchange as the server computes
the ECDHE shared secret. We made a list of 31 source-destination address pairs
for these indirect jumps, and constructed a \texttt{jump}/\texttt{ret} chain to
mimic the same jump pattern in our payload program. Our payload program mimics
the first 30 indirect jump source/destination pairs, with a final jump going to
a cache timing function in our payload program. However, due to the prior
pattern, this last jump is frequently mis-speculated (about 3.5\% of the time),
and instead goes to the destination corresponding to the 31st jump in OpenSSL,
which serves as our speculative entry point.

@@

%\FigSpectreOne

\if0
\subsection{Speculative Buffer Overflow Trigger}
\label{subsec:sbo-impl}

@@

We note significant unpredictability in the branch predictor, though we
show in Figure~\ref{fig:spectre-one} that a minimum around 200 safe input entries
to each dishonest entry are required before a distinguishable signal is present 
in the cache channel. 
\fi
%%%%%%%%%%%%%%
